Benchmarking of Rigetti architecture

### 2022-12-12 txt/rigetti.tex
creating appendix and rewriting mathemaytical prelimaties and parametrized family

Let us consider the problem of discrimination of von Neumann measurements 
$\PP_U$
and $\PP_{\Id}$ for $U = H_2 \diag(1, e^{i \phi}) H_2^\dagger$, 
	where $\phi \in [0, 2\pi)$. The schematic representation of theoretical 
	setup is shown in Fig.~\ref{fig:theoretical}. 
\begin{figure}[h!]
	\centering 
	\includegraphics[width=0.75\linewidth, angle=-90]{rys-theoretical.jpg}

	\caption{ A schematic representation of the setup for distinguishing
		measurements using entangled states. One of two known measurements 
		$\mathcal{P}_U$
		or $\PP_\Id$ is performed on part $A$ of the input state $\ket{\psi}$. 
		We use
		the output label $i$ and perform a conditional binary measurement
		$\mathcal{P}_{V_i}$ on part $B$. By the use of its output we formulate 
		our guess, 
		that is we decide  
		weather
		the measurement performed on part $A$ was $\mathcal{P}_U$ or 
		$\PP_\Id$. In out approach, if  $ j = 0$ we decide  $\PP_U$, if
		$j = 1$ we decide $\PP_{\Id}$. 
	}\label{fig:theoretical}
\end{figure} 
From ~\cite{puchala2018strategies}(Proposition 4) and Lemma~\ref{lemma:rho} we 
have 
	\begin{equation}
	\ket{\psi} = | \sqrt{\rho}^\top \rangle \rangle = \frac{1}{\sqrt{2}} |\Id_2 
	\rangle \rangle. 
	\end{equation}
	From Holevo-Helstrom theorem  we constrain a measurement $\mu$.  
	Let us define \begin{equation}
	X  = \left( \PP_U \otimes \Id_2 \right)(\proj{\psi}) -  \left( \PP_\Id 
	\otimes \Id_2 \right)(\proj{\psi})
	\end{equation}
	where $\ket{\psi}$ is defined in Remark~\ref{remark:discriminator}. From 
	Hahn-Jordan decomposition let \begin{equation}
	X = P - Q
	\end{equation}
	where $P, Q \ge 0 $. Observe, that $P $ and $Q$ are block-diagonal. 
	Let us define projectors $\Pi_P$ and $\Pi_Q$ onto  $im(P)$ and $im(Q)$, 
	respectively. Then  $\Pi_P$ and $\Pi_Q$ have the following forms
	\begin{equation}
	\Pi_P = \left(\begin{array}{cc}\proj{x_p}&0\\0&\proj{y_p}\end{array}\right) 
	\end{equation}
	and 
	\begin{equation}
	\Pi_Q = \left(\begin{array}{cc}\proj{x_q}&0\\0&\proj{y_q}\end{array}\right) 
	\end{equation}
	Hence, we define $V_0$ such that
	\begin{equation}
	\begin{split}
	V_0 \ket{x_p} = \ket{0} \\ 
	V_0 \ket{x_q} = \ket{1}
	\end{split}
	\end{equation}
	and $V_1$ such that
	\begin{equation}
	\begin{split}
	V_1 \ket{y_p} = \ket{0} \\ 
	V_1 \ket{y_q} = \ket{1}
	\end{split}
	\end{equation}
	The explicit form of $V_0$ and $V_1$ we can see in \texttt{one-qubit.nb}.
	Finally, we obtain the measurement $\mu$ of the form
	\begin{equation}
	\begin{split}
	 \mu(0) = \proj{0} \otimes V_0 \proj{0} V_0^\dagger +  \proj{1} \otimes V_1 
	 \proj{0} V_1^\dagger  \\ 
	\mu(1) = \proj{0} \otimes V_0 \proj{1} V_0^\dagger +  \proj{1} \otimes V_1 
	\proj{1} V_1^\dagger  
	\end{split}
	\end{equation}
\end{scheme}
This scheme is only theoretical. It is not possible to implement it on Rigetti 
architecture. For this, we will consider two approaches. The first one is 
described  by using postselection whereas the second one is considered by implementation 
of controlled unitary gate. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{figure}[h!]
	\centering 
	\includegraphics[width=0.75\linewidth, angle=-90]{rys-postsellection.jpg}

	\caption{ A schematic representation of the setup for distinguishing
		measurements using postselection. 
	}\label{fig:postsellection}
\end{figure}
\begin{enumerate}
\item We prepare input state $\ket{\psi} = \frac{1}{\sqrt{2} } | \Id_2 \rangle 
\rangle. $
\item We prepare one of two unitary channel $\Phi_{U} $ or $\Phi_{\1}$. 
\item We implement unitary $V_0 $ or $ V_1$.
\item We prepare the measurement $\Delta$ in standard basis (already exists on 
Rigetti architecture) on each qubits.
\item We calculate the probability of correct discrimination in the following 
way
\begin{equation}
p = ...
\end{equation}
\todo[inline]{grubsza kmina, musze Ci to na zywo opowiedziec xd}
\end{enumerate}
\end{scheme}

\newpage
\begin{scheme}(By using controlled unitary)
\begin{figure}[h!]
	\centering 
	\includegraphics[width=0.75\linewidth, angle=-90]{rys-controlled.jpg}

	\caption{ A schematic representation of the setup for distinguishing
		measurements using controlled unitary gate. 
	}\label{fig:controlled}
\end{figure}
\begin{enumerate}
\item We prepare input state $\ket{\psi} = \frac{1}{\sqrt{2} } | \Id_2 \rangle 
\rangle. $
\item We prepare one of two unitary channel $\Phi_{U} $ or $\Phi_{\1}$. 
\item We implement controlled unitary $V_0 \oplus V_1$.
\item We prepare the measurement $\Delta$ in standard basis (already exist on 
Rigetti architecture) on each qubits.
\item We calculate the probability of correct discrimination in the following 
way
\begin{equation}
p = \frac{|j=0 \wedge \Phi_{U} | + |j=1 \wedge \Phi_{\1}|}{\text{all 
cases}}
\end{equation}
\end{enumerate}
\end{scheme}

\section*{Acknowledgments}

>>>>

Let us consider the problem of discrimination of von Neumann measurements 
$\PP_U$
and $\PP_{\Id}$ for $U = H_2 \diag(1, e^{i \phi}) H_2^\dagger$, 
	where $\phi \in [0, 2\pi)$. The schematic representation of theoretical 
	setup is shown in Fig.~\ref{fig:theoretical}.

From ~\cite{puchala2018strategies}(Proposition 4) and Lemma~\ref{lemma:rho} we 
have 
	\begin{equation}
	\ket{\psi} = | \sqrt{\rho}^\top \rangle \rangle = \frac{1}{\sqrt{2}} |\Id_2 
	\rangle \rangle. 
	\end{equation}
	From Holevo-Helstrom theorem  we constrain a measurement $\mu$.  
	Let us define \begin{equation}
	X  = \left( \PP_U \otimes \Id_2 \right)(\proj{\psi}) -  \left( \PP_\Id 
	\otimes \Id_2 \right)(\proj{\psi})
	\end{equation}
	where $\ket{\psi}$ is defined in Remark~\ref{remark:discriminator}. From 
	Hahn-Jordan decomposition let \begin{equation}
	X = P - Q
	\end{equation}
	where $P, Q \ge 0 $. Observe, that $P $ and $Q$ are block-diagonal. 
	Let us define projectors $\Pi_P$ and $\Pi_Q$ onto  $im(P)$ and $im(Q)$, 
	respectively. Then  $\Pi_P$ and $\Pi_Q$ have the following forms
	\begin{equation}
	\Pi_P = \left(\begin{array}{cc}\proj{x_p}&0\\0&\proj{y_p}\end{array}\right) 
	\end{equation}
	and 
	\begin{equation}
	\Pi_Q = \left(\begin{array}{cc}\proj{x_q}&0\\0&\proj{y_q}\end{array}\right) 
	\end{equation}
	Hence, we define $V_0$ such that
	\begin{equation}
	\begin{split}
	V_0 \ket{x_p} = \ket{0} \\ 
	V_0 \ket{x_q} = \ket{1}
	\end{split}
	\end{equation}
	and $V_1$ such that
	\begin{equation}
	\begin{split}
	V_1 \ket{y_p} = \ket{0} \\ 
	V_1 \ket{y_q} = \ket{1}
	\end{split}
	\end{equation}
	The explicit form of $V_0$ and $V_1$ we can see in \texttt{one-qubit.nb}.
	Finally, we obtain the measurement $\mu$ of the form
	\begin{equation}
	\begin{split}
	 \mu(0) = \proj{0} \otimes V_0 \proj{0} V_0^\dagger +  \proj{1} \otimes V_1 
	 \proj{0} V_1^\dagger  \\ 
	\mu(1) = \proj{0} \otimes V_0 \proj{1} V_0^\dagger +  \proj{1} \otimes V_1 
	\proj{1} V_1^\dagger  
	\end{split}
	\end{equation}
\end{scheme}
This scheme is only theoretical. It is not possible to implement it on Rigetti 
architecture. For this, we will consider two approaches. The first one is 
described  by using postselection whereas the second one is considered by implementation 
of controlled unitary gate. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{enumerate}
\item We prepare input state $\ket{\psi} = \frac{1}{\sqrt{2} } | \Id_2 \rangle 
\rangle. $
\item We prepare one of two unitary channel $\Phi_{U} $ or $\Phi_{\1}$. 
\item We implement unitary $V_0 $ or $ V_1$.
\item We prepare the measurement $\Delta$ in standard basis (already exists on 
Rigetti architecture) on each qubits.
\item We calculate the probability of correct discrimination in the following 
way
\begin{equation}
p = ...
\end{equation}
\todo[inline]{grubsza kmina, musze Ci to na zywo opowiedziec xd}
\end{enumerate}
\end{scheme}

\newpage
\begin{scheme}(By using controlled unitary)

\begin{enumerate}
\item We prepare input state $\ket{\psi} = \frac{1}{\sqrt{2} } | \Id_2 \rangle 
\rangle. $
\item We prepare one of two unitary channel $\Phi_{U} $ or $\Phi_{\1}$. 
\item We implement controlled unitary $V_0 \oplus V_1$.
\item We prepare the measurement $\Delta$ in standard basis (already exist on 
Rigetti architecture) on each qubits.
\item We calculate the probability of correct discrimination in the following 
way
\begin{equation}
p = \frac{|j=0 \wedge \Phi_{U} | + |j=1 \wedge \Phi_{\1}|}{\text{all 
cases}}
\end{equation}
\end{enumerate}
\end{scheme}

PyQBench to biblioteka języka Python oraz narzędzie wiersza polecenia służące do testowania bramkowych komputerów kwantowych na podstawie ich zdolności do rozróżniania pomiędzy dwoma pomiarami von Neumanna w różnych bazach. Narzędzie wiersza polecenia pozwala na uruchomienie benchmarków wykorzystujących rozróżnianie między pomiarem w bazie obliczeniowej a pomiarem w bazie definiowanej przez sparametryzowaną rodzinę Fouriera. Użytkownik może wybierać, czy eksperyment zostanie przeprowadzony przy pomocy postselekcji czy alternatywnej metody “sumy prostej”, oraz kontrolować różne aspekty przeprowadzanego eksperymentu, takie jak liczba próbek używanych w samplowaniu, indeksy kubitów, zakresy kąta dla sparametryzowanej rodziny Fouriera. Jeżeli użytkownik chce użyć pomiaru w innej bazie, może wykorzystać PyQBench jako bibliotekę. Ten tryb użycia wymaga większego nakładu pracy, ale pozwala na rozszerzenie parametrów eksperymentu.  Warto wspomnieć, że w przypadku backendów udostępniających informacje o kalibracji kubitów, PyQBench wspiera mitygację błędów metodą matrix-free measurement mitigation. 
 Jednym z testowanych komputerów kwantowych był IBMQ, na którym wykonywane zostały eksperymenty na podstawie benchmarku rozróżniania pomiarów von Neumanna. 
Obecnie opracowujemy zebrane wyniki.
Biblioteka PyQBench będzie otwartym oprogramowaniem (ang. open source). Niebawem również zostanie wysłany artykuł do publikacji w czasopiśmie.  
\section*{Acknowledgments}
